---
title: Final Exam
author: Katy Linich
date: 5 May 2020
output: html_notebook
---

```{r}
library(tidyverse)

demographics <- read_csv("Job Prep Demographics.csv",
                         col_types = cols("Program Type"=col_character()))
demographics <- demographics %>%
  fill(`Program Type`, .direction = c("down"))

gpa <- read_csv("Job Prep GPA Data.csv", skip = 4)
survey <- read_csv("Job Prep Survey Results.csv", skip = 3)


job_data <- demographics %>%
  left_join(gpa, by="ID") %>%
  left_join(survey, by="ID")

job_data <- job_data %>%
  mutate(avg_GPA=((job_data$`S1 GPA`+job_data$`S2 GPA`)/2))
```

Here is my previous script for reversed likert items:

```{r}
library(here)
total_scores <- function(survey_data, scale_num, vector_num) {
  survey_matrix <- as.matrix(survey_data)
  survey_matrix[,vector_num] <- (scale_num+1) - survey_matrix[,vector_num]
  sum_total_scores <- rowSums(survey_matrix)
  survey_data$sum <- sum_total_scores
  return(survey_data)
}
```


Here is my previous script for residuals:
```{r}
library(here)
largest_residuals <- function(x,y,p) { 
 linear_model <- lm(y~x)
 residual <- abs(resid(linear_model))
 sort_resid <- sort(residual, decreasing = TRUE)
 low_resid <- sort_resid[p]
 final_resid <- (residual>=low_resid)
  return(final_resid)
}
```


First, I'm using my equation to compile the scores of the assessment.
```{r}
job_scores <- total_scores(survey[,2:15], 6, c(4,9,12,13))

all_job_data <- job_data %>%
  left_join(job_scores)

```

Next, I'm going to create a plot to show the residuals in the data.
```{r}
gpa_resid <- largest_residuals(all_job_data$sum, all_job_data$avg_GPA)

gpa_plot <- plot(all_job_data$sum~all_job_data$avg_GPA,
     main = "Relationship between Avg GPA and Optimism",
     xlab = "Avg GPA",
     ylab = "Optimism",
     col = ifelse(gpa_resid, "red","black"))
```
This could be an issue, because there's no real discernable trend when you use average GPA. It may be more useful to use weighted GPA.




Now I'm going to create a column for weighted GPA.
```{r}
all_job_data <- all_job_data %>%
  mutate(weighted.gpa=(((all_job_data$`S1 Credits`*all_job_data$`S1 GPA`)+(all_job_data$`S2 Credits`*all_job_data$`S2 GPA`))/(all_job_data$`S1 Credits`+all_job_data$`S2 Credits`)))
```

Now we can work on answering the research questions.

Question 1: Is senior-year grade point average (GPA) related to the degree of optimism about future employment?

```{r}
ggplot(all_job_data, aes(all_job_data$weighted.gpa, all_job_data$sum)) +
  geom_point() + geom_smooth(method = "lm", se=FALSE) + xlab("Weighted GPA") +
  ylab("Level of Optimism") + labs(title="Relationship between Senior Year Weighted GPA and Optimism About Employment")
```
From what we can see here, there appears to be a very slight, negative relationship between weighted GPA and level of optimism about future employment. 


Question 2: Is the relationship of GPA and optimism consistent or different for different categories of disciplines?

```{r}
ggplot(all_job_data, aes(all_job_data$weighted.gpa, all_job_data$sum, color=all_job_data$`Program Type`)) +
  geom_point() + geom_smooth(method = "lm", se=FALSE) + 
  xlab("Weighted GPA") +
  ylab("Level of Optimism") + labs(title="Relationship between Senior Year Weighted GPA and Optimism About Employment", color="Program Type")
```


